Statistical Inference for inter-arrival times of extreme events in bursty time series

### 2020-01-16 article_springer/article_springer.Rmd
Update ... submitted version

authors: 
- name: "Katharina Hees"
  address: TU Dortmund University
  email: hees@statistik.tu-dortmund.de

- name: "Smarak Nayak"
  address: National Australia Bank
  email: smarak.nayak@nab.com.au

- name: "Peter Straka*"
  address: UNSW Sydney
  email: p.straka@unsw.edu.au

@@

  This article proposes a new model for the _inter-exceedance times_ of thresholds;
  the threshold exceedances are modeled via the standard 
  Peaks-Over-Threshold (POT) method.
  For high thresholds and 
  infinite-mean waiting times, we show that the times between threshold 
  crossings are Mittag-Leffler distributed, and thus form a "fractional 
  Poisson Process" which generalizes the standard Poisson Process of threshold
  exceedances.
  We provide graphical means of estimating model parameters and assessing 
  model fit. Along the way, we apply our inference method to a real-world
  bursty time series, and 
  show how the memory of the Mittag-Leffler distribution affects 
  the predictive distribution for the time until the next extreme event.

bibliography: CTRMstats.bib
output: 
       rticles::elsevier_article:
                fig_caption: yes
                number_sections: true

@@

```{r get-rticles, eval=FALSE, include=FALSE, cache=TRUE}
# use this rticles version: 
devtools::install_github("rstudio/rticles")
# use the newest MittagLeffleR version:
devtools::install_github("UNSW-MATH/MittagLeffleR")
library(MittagLeffleR)
devtools::install_github("strakaps/CTRE")
library("CTRE")
library(CTRE)
install.packages("here")
library(here)
Sys.setenv(TEXINPUTS=getwd(),
           BIBINPUTS=getwd(),
           BSTINPUTS=getwd())
```

@@

Time series displaying temporally inhomogeneous behaviour have received 
strong interest in the recent statistical physics literature 
[@Barabasi2005; @Oliveira2005; @Vasquez2006; @Vazquez2007; @Omi2011; 
@Min2010; @Karsai2011; @Bagrow2013]. They have been observed in the context 
of earthquakes, sunspots, neuronal activity and human communication [see 
@Karsai2012; @Vajna2013; @MeerschaertStoev08 for a list of references]. 
Such time series exhibit high activity in some 'bursty' intervals, which 
alternate with other, quiet intervals.  Although several mechanisms are 
plausible explanations for bursty behaviour (most prominently self-exciting
point process by @hawkes1971point), there seems to be one salient feature 
which very typically indicates the departure from temporal homogeneity: a 
heavy-tailed distribution of waiting times [@Vasquez2006; @Karsai2012; 
@Vajna2013]. As we show below in simulations, a simple renewal process with
heavy-tailed waiting times can capture this type of dynamics. For many 
systems, the renewal property is appropriate; a simple test of the absence 
of correlations in a succession of waiting times can be undertaken by 
randomly reshuffling the waiting times [@Karsai2012].

@@

Maxima of events which occur according to a renewal process with
heavy-tailed waiting times have been studied under 
the names "Continuous Time Random Maxima process" (CTRM) 
[@Benson2007; @MeerschaertStoev08; @Hees16; @Hees17], 
"Max-Renewal process" [@Silvestrov2002a; @ST04; @Basrak2014], 
and "Shock process"
[@Esary1973; @Sumita1983; @Sumita1984; @Sumita1985; @Anderson1987; @Gut1999].
The existing literature focuses on probabilistic results surrounding these 
models. 
In this work, however, we introduce a method of inference for this type of 
model, which is seemingly not available in the literature.

We review the marked renewal process in Section 2, and 
derive a scaling limit theorem for inter-exceedance times in Section 3.
We give a statistical procedure to estimate model parameters via stability plots 
in Section 5, but to set the stage we first discuss inference for the 
Mittag-Leffler distribution in Section 4. In Section 6, we discuss the memory property of the 
Mittag-Leffler distribution, and how it affects the predictive distribution 
for the time until the next threshold crossing event. A simulation study of the effectiveness of our statistical procedure is given in Section 7. In Section 8 we apply our method to a real data set. Finally we close with a discussion and conclusion in Section 8.
For all statistical computations we have used R [@R]. 
All code and data used for the analysis in this article has been organized into 
an R package `CTRE` (<https://github.com/strakaps/CTRE>).
The source code for the figures generated in this manuscript is 
available online at <https://github.com/strakaps/bursty-POT>.

@@

```{r thresholdedBursty, message=FALSE, fig.height=6, fig.width=7,out.width='70%',fig.align='center',fig.cap="\\label{fig:thresholdedBursty} Exceedances (red) and times until Exceedance (durations between blue crosses) for a given threshold $\\ell$ (dashed line)."}
tail <- params$tail
n <- params$n 
sigma <- (cos(pi*tail/2))^(1/tail)
times <- cumsum(stabledist::rstable(n, tail, 1, sigma, pm=1))
magnitudes <- extRemes::revd(n, scale = 1, shape = 0)
sim_ctre <- ctre(data.frame(times, magnitudes)) 
p1<-plot(sim_ctre, p = 0.01, main = "Simulated MRP") 
p2<-flares %>% ctre() %>% plot(p = 0.02, log = 'y', main = "HXRBS data") 
grid.arrange(p1,p2)

@@

In this article, we restrict ourselves to the uncoupled case, where $W$ and $J$ are independent. Then the two sequences $X(\ell, n)_{n \in \mathbb N}$ and $T(\ell, n)_{n \in \mathbb N}$ are independent as well. To see why, note that $X(\ell)$ is, in distribution, simply equal to $J | J > \ell$, independent of any waiting time $W_k$.  
Figure \ref{fig:thresholdedBursty} shows a simulated dataset in the top panel, where $W$ has a stable distribution with tail parameter $\beta =$ `r params$tail` (and skewness $1$ and location $0$), and where $J$ is from a standard Gumbel distribution. In the bottom panel, we plot a time series of solar flare intensities derived from a NASA dataset [@HXRBS] which we will later examine more closely (see Section 8). Clearly, the simulated data exhibit long intervals 
_without any_ events, whereas in the real-world dataset events appear continuously. The threshold exceedances, however, appear to have visually similar statistical behaviour in both models. Observations below a threshold are commonly discarded in Extreme Value Theory (POT approach); likewise, the CTRE model interprets these observations as noise and discards them.

@@

Since the Mittag-Leffler distribution is heavy-tailed, many researchers would intuitively give the highest importance to the tail behaviour of the distribution. Of course, one can also use established tail exponent estimators for the estimation of the parameter $\beta$ such as the Hill estimator [@hill1975simple]
\begin{align}\label{eq:hill}
H_{k,n}= \frac{1}{k} \sum_{i=0}^{k-1} \log \frac{X_{n-i,n}}{X_{n-k,n}}. 
\end{align}
However, these methods are statistically less efficient since they only use a portion of the information contained in the data, as mentioned by @kozubowski2001. Moreover, the hill estimator requires a tuning parameter, and only performs reliably well for distributions close to Pareto [see @Resnick97]. Figure \ref{fig:Hillplots} shows Hill plots for Mittag-Leffler simulated data, with varying sample sizes and tails. To deduce the correct tail parameter estimates from these plots is virtually impossible. Furthermore the Hill estimator of course completely fails if the inter-arrival times are exponential distributed and not heavy-tailed.

```{r Hillplots, echo=FALSE,warning=FALSE, fig.align="center", fig.cap="\\label{fig:Hillplots} Hillplots for simulated Mittag-Leffler data with true tail 0.8 and sample size 200 (left) and 1000 (right), with number of upper order statics k on which the hill estimator is based on the x-axis. ", out.width = '90%'}
hill_estim<-function(k,data){
  n<-length(data)
  data<-sort(data)
  data_thinned<-data[(n-k):n]
  r<-1/k * sum(log(data_thinned/data_thinned[1]))
  return(1/r)
}

@@

```{r LRT_power, echo=FALSE,warning=FALSE, fig.align="center", fig.cap="\\label{Fig:LRT} Power for bootstrapped LRT for different sample sizes , varying tails and scale parameter equal to 1.",out.width='90%'}
load(here::here("article_springer","article_springer_files","ML_LRT_simu","estimates_df_LRT_boot1_01.RData"))
df_LRT_boot_01 %>% filter(n=="20" || n=="50" || n=="100" || n=="200") %>% ggplot(aes(x=true.tail,y=power)) +  geom_line(aes(y=power,group=factor(n),color=factor(n),linetype=factor(n)),size=0.7)+
  scale_linetype_manual(values = c("solid","dashed","twodash","dotted"),name  ="Sample size")+
   labs(y="Power",x= "True tail parameter")+
  scale_color_manual(values=c("red", "blue","darkgreen","black"), name  ="Sample size")
```

@@

To test our inference method via stability plots, we have simulated 
independent waiting time and magnitude pairs $(W_k, J_k)$ for waiting times that follow

@@

(iii) an Exponential distribution.

In order to have exact analytical values available for $\beta$ and $\sigma_0$, a distribution for $W_k$ needs to be chosen for which $b(n)$ from \eqref{eq:stability} is known. For (i) we choose $W_k \stackrel{d}{=} D$, where $D$ is as in \eqref{eq:stability}, then due to the stability property we have the *equality* of distribution
$W_1 + \ldots + W_n \stackrel{d}{=} b(n) D$, 
for $b(n) = n^{1/\beta}$. 
Using the parametrisation of @SamorodnitskyTaqqu, a few lines of 
calculation [see e.g. the vignette on parametrisation in @MittagLeffleR]
show that $D$ must have the stable distribution 
$S_\beta(\cos(\pi \beta/2)^{1/\beta}, +1, 0)$, which is 
implemented in the R package `stabledist` by @stabledist. By the Theorem, the distribution of $T(\ell)$ is approximately 
$$
{\rm ML}(\beta, p^{-1/\beta}) 
= {\rm ML}(\beta, k^{-1/\beta} n^{1/\beta}),
$$
which means 
$\sigma_0 = n^{1/\beta}$. In the Pareto example we choose $P(W>t)=Ct^{-\beta}$ with $C=(1/\Gamma(1-\beta))^{1/\beta}$. We have chosen $\beta=0.8$ in the stable as well as in the Pareto case. As third example we choose Exponential distributed waiting times with a rate of $1$. Figure \ref{Fig:TailSimu} shows the stability plots for the estimation of the tail parameter. Therefore, $\hat \beta(k)$ vs. $k$ is plotted for each of the $m=100$ simulation runs (grey thin lines). In the left column one can see the estimates based on the log-moment estimator, the stability plots in the middle are based on the maximum-likelihood estimator and on the right hand on the hill estimator. Recall that $k$ is the index of the order statistics of $J_k$ at which the threshold $\ell$ is placed. Note, that the hill estimator is here based on the waiting times, not on the inter-exceedance times, hence the estimation is based on much more data. If one would base the hill-estimation only on the inter-exceedance times, the sample size would be too small and it would be impossible to deduce an estimate for the tail parameter from the plots (compare Figure \ref{fig:Hillplots}). The $k$ on the $x-$axis is in the plots for the hill estimator the $k$ which defines on which orders stats the hill estimator is based on (compare \eqref{eq:hill}). The bigger sample size results in the middle in a lower bias. But, one can also see, that there is much more variance between the stability plots for the different simulation runs in case of the hill estimator compared to the other two estimators. The black line is the mean value of the point estimators for each $k$. However, if one wants to get a single estimate via such a stability plot, it would be much more difficult to deduce it from a hill plot compared to the other two estimators. In case that one has only thresholded data and hence a much smaller simple size, the hill estimator gets useless (see Section \ref{sec:ML}). Another advantage of the other estimators compared to the hill estimator is, that in the case one has exponential waiting times, the log-moment as well as the mle can detect this by getting an estimate round about $1$. The hill estimators is just a tail parameter estimator for Pareto like tails and hence of course fails in case of Exponential inter-exceedance times. Figure \ref{Fig:ScaleSimu} shows the stability plots for the different estimators for the scale parameter. Here of course only for the log-moment (on the right hand)  and the mle (on the left hand), since the hill estimator is just a tail parameter estimate. Here again both estimators show good performance. The stability plots in the mle case are a little bit closer to the true value, therefore, the mle is much more computational intesive.

```{r TailSimuplots, message=FALSE,echo=FALSE,fig.height=7,fig.align="center",fig.cap="\\label{Fig:TailSimu} Stability Plots for m=100 simulation runs for stable distributed waiting times with a tail parameter of 0.8 (top row), Pareto distributed waiting times with tail parameter 0.8 (middle row) and Exponential distributed waiting times (lower row). Left column: log-moment estimator, middle column: MLE, right column: Hill estimator. The grey thin lines are the stability plots for the different simulation runs and the dark lines are their means. The red dotted line shows the true tail parameter. ", fig.show='hold', out.width = '100%'}

@@

#Hill estimator
hill_Ex1_tail<-ggplot()+ geom_line(data=tbl_estimates_tail08_hill_Ex1,aes(x=k,y=value,group=rep),colour="lightgrey") +
  coord_cartesian(ylim = c(0.5, 1.5)) +
  geom_line(data=means.tbl_hill_Ex1,aes(x=k, y=mw),colour="black") + ylab("Estimated tail") +  geom_hline(yintercept=0.8, linetype="dashed", color = "red")

@@

#Hill estimator
hill_Ex2_tail<-ggplot() + geom_line(aes(x=k,y=value,group=rep),colour="lightgrey",data=tbl_estimates_Ex2_Pareto_tail08_hill_times) +
  coord_cartesian(ylim = c(0.5, 1.5)) +
  geom_line(aes(x=k, y=mw),colour="black",data=means.tbl_hill_times_Ex2) + ylab("Estimated tail") +  geom_hline(yintercept=0.8, linetype="dashed", color = "red")

@@

#Hill estimator
hill_Ex3_tail<-ggplot() + geom_line(aes(x=k,y=value,group=rep),colour="lightgrey",data=tbl_estimates_Example3_Exp_hill_times) +
  coord_cartesian(ylim = c(0.5, 10)) +
  geom_line(aes(x=k, y=mw),colour="black",data=means.tbl_hill_times_Ex3) + ylab("Estimated tail") + geom_hline(yintercept=1, linetype="dashed", color = "red")

@@

```{r ScaleSimuplots, message=FALSE,echo=FALSE,fig.height=7,fig.align="center",fig.cap="\\label{Fig:ScaleSimu} Stability Plots for m=100 simulation runs for stable distributed waiting times with a tail parameter of 0.8 (top row), Pareto distributed waiting times with tail parameter 0.8 (middle row) and Exponential distributed waiting times (lower row). Left column: log-moment estimator, right column: MLE. The grey thin lines are the stability plots for the different simulation runs and the dark lines are their means. The red dotted line shows the true tail parameter.", fig.show='hold', out.width = '90%'}
#Scale Parameter

@@

We now want to apply the proposed method to a real data example, the solar flare data which was already mentioned in Section 1 and can be seen in Figure \ref{fig:thresholdedBursty}. The data were extracted from the "complete Hard X Ray Burst Spectrometer event list", a comprehensive reference for all measurements of the Hard X Ray Burst Spectrometer on NASA's Solar Maximum Mission from the time of launch on Feb 14, 1980 to the end of the mission in Dec 1989. 12,776 events were detected, with the "vast  majority being solar flares". The list includes the start time, peak time, duration, and peak rate of each event. We have used "start time" as the variable for event times, and "peak rate" as the variable for event magnitudes.

@@

```{r flare-diagnostics-2, fig.height=4.5, fig.width=4.5, fig.align="center",fig.cap="\\label{fig:flare-diagnostics-2} Diagnostic plots for the solar flare data: empirical copula and QQ Plot.",out.width="70%"}
empcopula(flares_ctre)
```

```{r flare-diagnostics-3, fig.height=4.5, fig.width=4.5, fig.align="center",fig.cap="\\label{fig:flare-diagnostics-3} Diagnostic plots for the solar flare data: empirical copula and QQ Plot.",out.width="70%"}
flares_ctre %>% interarrival() %>% mlqqplot(log = 'xy', tail = 0.85, main = "Mittag-Leffler QQ Plot")
```

Figure \ref{fig:flares} shows the stability plots for the solar flare data, on the left for the tail parameter and on the right for the scale parameter. Dotted lines show 95% confidence intervals, which are derived from the asymptotic normality of the log-moments estimators [@Cahoy2013] and the $\delta$-method [@MittagLeffleR, dashed lines show the actual values of $\beta$ resp.\ $\sigma_0$. The stability plot for the tail stabilizes nicely round about 0.85 (dotted line), while it is a little bit harder to deduce an estimate for the scale parameter. Howerever, the stability plot for the scale parameter seems to stabilize round about $3 \times 10^7$.

>>>>

author:

- name: "Katharina Hees"
  affiliation: a
  email: "hees@statistik.tu-dortmund.de"

- name: "Smarak Nayak"
  affiliation: b
  email: "smarak.nayak@nab.com.au"

- name: "Peter Straka"
  affiliation: c
  email: "p.straka@unsw.edu.au"

address:

- code: a
  address: "Department of Statistics, TU Dortmund University, Dortmund, Germany"

- code: b
  address: "National Australia Bank, Melbourne, Australia"

- code: c
  address: "School of Mathematics and Statistics, UNSW, Sydney, Australia"

corrauth:
  name: "Katharina Hees"
  address: "Department of Statistics, TU Dortmund University, Dortmund, Germany"

@@

  This article proposes a new model for the _inter-exceedance times_ of events 
  above high thresholds; the threshold exceedances itself are modeled via the standard 
  Peaks-Over-Threshold (POT) method. For high thresholds and 
  infinite-mean waiting times, we show that the times between threshold 
  crossings are Mittag-Leffler distributed, and thus form a "fractional 
  Poisson Process" which generalizes the standard Poisson Process of threshold
  exceedances.
  We provide graphical means of estimating model parameters and assessing 
  model fit. Along the way, we apply our inference method to a real-world
  bursty time series, and 
  show how the memory of the Mittag-Leffler distribution affects 
  the predictive distribution for the time until the next extreme event.

bibliography: CTRMstats.bib
csl: 
output: 
       rticles::elsevier_article:
                fig_caption: yes
                number_sections: true

@@

```{r get-rticles, eval=FALSE, include=FALSE, cache=TRUE}
# use this rticles version: 
devtools::install_github("rstudio/rticles")
# use the newest MittagLeffleR version:
devtools::install_github("strakaps/MittagLeffleR")
library(MittagLeffleR)
devtools::install_github("strakaps/CTRE")
library("CTRE")
library(CTRE)
install.packages("here")
library(here)
Sys.setenv(TEXINPUTS=getwd(),
           BIBINPUTS=getwd(),
           BSTINPUTS=getwd())
```

@@

Time series displaying temporally inhomogeneous behaviour in terms of the occurrence of events have received strong interest in the recent statistical physics literature [@Barabasi2005;@Oliveira2005; @Vasquez2006; @Vazquez2007; @Omi2011; @Min2010; @Karsai2011;@Bagrow2013]. They have been observed in the context of earthquakes, sunspots, neuronal activity and human communication [see @Karsai2012; @Vajna2013; @MeerschaertStoev08 for a list of references]. 
Such time series exhibit high activity in some 'bursty' intervals, which 
alternate with other, quiet intervals.  Although several mechanisms are 
plausible explanations for bursty behaviour (most prominently self-exciting
point process by @hawkes1971point), there seems to be one salient feature 
which very typically indicates the departure from temporal homogeneity: a 
heavy-tailed distribution of waiting times [@Vasquez2006; @Karsai2012; 
@Vajna2013]. As we show below in simulations, a simple renewal process with
heavy-tailed waiting times can capture this type of dynamics. For many 
systems, the renewal property is appropriate; a simple test of the absence 
of correlations in a succession of waiting times can be undertaken by 
randomly reshuffling the waiting times [@Karsai2012].

@@

Maxima of events which occur according to a renewal process with
heavy-tailed waiting times have been studied under 
the names "Continuous Time Random Maxima process" (CTRM) 
[@Benson2007; @MeerschaertStoev08; @hees2016joint; @hees2017coupled], 
"Max-Renewal process" [@Silvestrov2002a; @ST04; @Basrak2014], 
and "Shock process"
[@Esary1973; @Sumita1983; @Sumita1984; @Sumita1985; @Anderson1987; @Gut1999].
The existing literature focuses on probabilistic results surrounding these 
models. 
In this work, however, we introduce a method of inference for this type of 
model, which is seemingly not available in the literature.

We review the marked renewal process in Section 2, and 
derive a scaling limit theorem for inter-exceedance times in Section 3.
We give a statistical procedure to estimate model parameters via stability plots 
in Section 5, but to set the stage we first discuss inference for the 
Mittag-Leffler distribution in Section 4. A simulation study of the effectiveness of our statistical procedure is given in Section 6. In Section 7 we apply our method to a real data set.  In Section 8, we discuss the memory property of the  Mittag-Leffler distribution, and how it affects the predictive distribution for the time until the next threshold crossing event. Finally we close with a discussion and conclusion in Section 9.
For all statistical computations we have used R [@R]. 
All code and data used for the analysis in this article has been organized into 
an R package `CTRE` (<https://github.com/strakaps/CTRE>).
The source code for the figures generated in this manuscript is 
available online at <https://github.com/strakaps/bursty-POT>.

@@

```{r thresholdedBursty, message=FALSE, fig.height=6, fig.width=7,out.width='70%',fig.align='center',fig.cap="\\label{fig:thresholdedBursty} Exceedances (red) and times until Exceedance (durations between blue crosses) for a given threshold $\\ell$ (dashed line). Upper picture: Simulated data with stable distributed waiting times. Lower picture: Solar flares during 1982."}
tail <- params$tail
n <- params$n 
sigma <- (cos(pi*tail/2))^(1/tail)
times <- cumsum(stabledist::rstable(n, tail, 1, sigma, pm=1))
magnitudes <- extRemes::revd(n, scale = 1, shape = 0)
sim_ctre <- ctre(data.frame(times, magnitudes)) 
p1<-plot(sim_ctre, p = 0.01, main = "Simulated MRP") 
p2<-flares %>% ctre() %>% plot(p = 0.02, log = 'y', main = "HXRBS data") 
grid.arrange(p1,p2)

@@

In this article, we restrict ourselves to the uncoupled case, where $W$ and $J$ are independent. Then the two sequences $X(\ell, n)_{n \in \mathbb N}$ and $T(\ell, n)_{n \in \mathbb N}$ are independent as well. To see why, note that $X(\ell)$ is, in distribution, simply equal to $J | J > \ell$, independent of any waiting time $W_k$.  
Figure \ref{fig:thresholdedBursty} shows a simulated dataset in the top panel, where $W$ has a stable distribution with tail parameter $\beta =$ `r params$tail` (and skewness $1$ and location $0$), and where $J$ is from a standard Gumbel distribution. In the bottom panel, we plot a time series of solar flare intensities derived from a NASA dataset [@HXRBS] which we will later examine more closely (see Section 7). Clearly, the simulated data exhibit long intervals 
_without any_ events, whereas in the real-world dataset events appear continuously. The threshold exceedances, however, appear to have visually similar statistical behaviour in both models. Observations below a threshold are commonly discarded in Extreme Value Theory (POT approach); likewise, the CTRE model interprets these observations as noise and discards them.

@@

Since the Mittag-Leffler distribution is heavy-tailed, many researchers would intuitively give the highest importance to the tail behaviour of the distribution. Of course, one can also use established tail exponent estimators for the estimation of the parameter $\beta$ such as the Hill estimator [@hill1975simple] based on the $r+1$ upper order statistics
\begin{align}\label{eq:hill}
H_{r,n}=\left[ \frac{1}{r} \sum_{i=0}^{r} \log \frac{X_{(i)}}{X_{(r+1)}}\right]^{-1}, 
\end{align}
where $X_{(1)} \geq X_{(2)} \geq \ldots \geq X_{(n)}$ denote the order statistics in decreasing order of a sequence $X_1,\ldots,X_n$. However, these methods are statistically less efficient since they only use a portion of the information contained in the data, as mentioned by @kozubowski2001. Moreover, the hill estimator requires a tuning parameter, denoted with $r$ in \eqref{eq:hill}. Additionaly, the estimator only performs reliably well for distributions close to Pareto [see @Resnick97]. Figure \ref{fig:Hillplots} shows Hill plots for Mittag-Leffler simulated data, with varying sample sizes and tails. To deduce the correct tail parameter estimates from these plots is virtually impossible. To estimate the tail parameter of a Mittag-Leffler distribution with the hill estimator, the sample size has to be much larger and one has to choose a large tuning parameter $r$. In case that only events above a high threshold are recorded, even $n=1000$ events are unrealistic. Furthermore the Hill estimator of course completely fails if the inter-arrival times are exponential distributed and not heavy-tailed. We will comeback to the problems with the hill estimtator in Section \ref{Simulationstudy}.

```{r Hillplots, echo=FALSE,warning=FALSE, fig.align="center", fig.cap="\\label{fig:Hillplots} Hillplots for simulated Mittag-Leffler data with true tail 0.8 and sample size 200 (left) and 1000 (right), with number of upper order statics r on which the hill estimator is based on the x-axis. ", out.width = '90%'}
hill_estim<-function(k,data){
  n<-length(data)
  data<-sort(data)
  data_thinned<-data[(n-k):n]
  r<-1/k * sum(log(data_thinned/data_thinned[1]))
  return(1/r)
}

@@

```{r LRT_power, echo=FALSE,warning=FALSE, fig.align="center", fig.cap="\\label{Fig:LRT} Power for bootstrapped LRT for different sample sizes , varying tails and scale parameter equal to 1.",out.width='90%'}
load(here::here("article_springer","article_springer_files","ML_LRT_simu","estimates_df_LRT_boot1_01.RData"))
df_LRT_boot_01 %>% filter(n=="20" || n=="50" || n=="100" || n=="200") %>% ggplot(aes(x=true.tail,y=power)) +  geom_line(aes(y=power,group=factor(n),color=factor(n),linetype=factor(n)),size=0.5)+
  scale_linetype_manual(values = c("solid","dashed","twodash","dotted"),name  ="Sample size")+
   labs(y="Power",x= "True tail parameter")+
  scale_color_manual(values=c("red", "blue","darkgreen","black"), name  ="Sample size")
```

@@

To test our inference method via stability plots, we have simulated $m=100$ times $n=10000$ independent waiting time and magnitude pairs $(W_k, J_k)$ for waiting times that follow

@@

(iii) an exponential distribution.

In order to have exact analytical values available for $\beta$ and $\sigma_0$, a distribution for $W_k$ needs to be chosen for which $b(n)$ from \eqref{eq:stability} is known. For (i) we choose $W_k \stackrel{d}{=} D$, where $D$ is as in \eqref{eq:stability}, then due to the stability property we have the *equality* of distribution
$W_1 + \ldots + W_n \stackrel{d}{=} b(n) D$, 
for $b(n) = n^{1/\beta}$. 
Using the parametrisation of @SamorodnitskyTaqqu, a few lines of 
calculation [see e.g. the vignette on parametrisation in @MittagLeffleR]
show that $D$ must have the stable distribution 
$S_\beta(\cos(\pi \beta/2)^{1/\beta}, +1, 0)$, which is 
implemented in the R package `stabledist` by @stabledist. By the Theorem, the distribution of $T(\ell)$ is approximately 
$$
{\rm ML}(\beta, p^{-1/\beta}) 
= {\rm ML}(\beta, k^{-1/\beta} n^{1/\beta}),
$$
which means 
$\sigma_0 = n^{1/\beta}$. In the Pareto example we choose $P(W>t)=Ct^{-\beta}$ with $C=(1/\Gamma(1-\beta))^{1/\beta}$. We have chosen $\beta=0.8$ in the stable as well as in the Pareto case. As third example we choose exponential distributed waiting times with a rate of $1$. Figure \ref{Fig:TailSimu} shows the stability plots for the estimation of the tail parameter. Therefore, $\hat \beta(k)$ vs. $k$ is plotted for each of the $m=100$ simulation runs (grey thin lines). In the left column one can see the estimates based on the log-moment estimator, the stability plots in the middle are based on the maximum-likelihood estimator and on the right hand on the hill estimator. Recall that $k$ is the index of the order statistics of $J_k$ at which the threshold $\ell$ is placed. Note, that the hill estimator is here based on the waiting times, not on the inter-exceedance times, hence the estimation is based on much more data. If one would base the hill-estimation only on the inter-exceedance times, the sample size would be too small and it would be impossible to deduce an estimate for the tail parameter from the plots (compare Figure \ref{fig:Hillplots}). The $k$ on the $x-$axis is in the plots for the hill estimator the $r$ which defines on which upper orders stats the hill estimator is based on (compare \eqref{eq:hill}). The bigger sample size results in the middle in a lower bias. But, one can also see, that there is much more variance between the stability plots for the different simulation runs in case of the hill estimator compared to the other two estimators. The black line is the mean value of the point estimators for each $k$ resp. $r$. However, if one wants to get a single estimate via such a stability plot, it would be much more difficult to deduce it from a hill plot compared to the other two estimators. In case that one has only thresholded data and hence a much smaller simple size, the hill estimator gets useless (see Section \ref{sec:ML}). Another advantage of the other estimators compared to the hill estimator is, that in the case one has exponential waiting times, the log-moment as well as the mle can detect this by getting an estimate close to $1$. The hill estimators is just a tail parameter estimator for Pareto like tails and hence of course fails in case of exponential distributed inter-exceedance times. Figure \ref{Fig:ScaleSimu} shows the stability plots for the different estimators for the scale parameter. Here of course only for the log-moment (on the right hand)  and the mle (on the left hand), since the hill estimator is just a tail parameter estimate. Here again both estimators show good performance. The stability plots in the mle case are a little bit closer to the true value, except in the exponetial case. Therefore, the mle is much more computational intensive.

```{r TailSimuplots, message=FALSE,echo=FALSE,fig.height=7,fig.align="center",fig.cap="\\label{Fig:TailSimu} Stability Plots for m=100 simulation runs for stable distributed waiting times with a tail parameter of 0.8 (top row), Pareto distributed waiting times with tail parameter 0.8 (middle row) and exponential distributed waiting times (lower row). Left column: log-moment estimator, middle column: MLE, right column: Hill estimator. The grey thin lines are the stability plots for the different simulation runs and the dark lines are their means. The red dotted line shows the true tail parameter. ", fig.show='hold', out.width = '100%'}

@@

#Hill estimator
hill_Ex1_tail<-ggplot()+ geom_line(data=tbl_estimates_tail08_hill_Ex1,aes(x=k,y=value,group=rep),colour="lightgrey") +
  coord_cartesian(ylim = c(0.5, 1.5)) +
  geom_line(data=means.tbl_hill_Ex1,aes(x=k, y=mw),colour="black") + ylab("Estimated tail") +  geom_hline(yintercept=0.8, linetype="dashed", color = "red")+ xlab("r")

@@

#Hill estimator
hill_Ex2_tail<-ggplot() + geom_line(aes(x=k,y=value,group=rep),colour="lightgrey",data=tbl_estimates_Ex2_Pareto_tail08_hill_times) +
  coord_cartesian(ylim = c(0.5, 1.5)) +
  geom_line(aes(x=k, y=mw),colour="black",data=means.tbl_hill_times_Ex2) + ylab("Estimated tail") +  geom_hline(yintercept=0.8, linetype="dashed", color = "red")+ xlab("r")

@@

#Hill estimator
hill_Ex3_tail<-ggplot() + geom_line(aes(x=k,y=value,group=rep),colour="lightgrey",data=tbl_estimates_Example3_Exp_hill_times) +
  coord_cartesian(ylim = c(0.5, 10)) +
  geom_line(aes(x=k, y=mw),colour="black",data=means.tbl_hill_times_Ex3) + ylab("Estimated tail") + geom_hline(yintercept=1, linetype="dashed", color = "red")+ xlab("r")

@@

```{r ScaleSimuplots, message=FALSE,echo=FALSE,fig.height=7,fig.align="center",fig.cap="\\label{Fig:ScaleSimu} Stability Plots for m=100 simulation runs for stable distributed waiting times with a tail parameter of 0.8 (top row), Pareto distributed waiting times with tail parameter 0.8 (middle row) and exponential distributed waiting times (lower row). Left column: log-moment estimator, right column: MLE. The grey thin lines are the stability plots for the different simulation runs and the dark lines are their means. The red dotted line shows the true tail parameter.", fig.show='hold', out.width = '90%'}
#Scale Parameter

@@

We now want to apply the proposed method to a real data example, the solar flare data which was already mentioned in Section 1 and can be seen in Figure \ref{fig:thresholdedBursty}. The data were extracted from the "complete Hard X Ray Burst Spectrometer event list", a comprehensive reference for all measurements of the Hard X Ray Burst Spectrometer on NASA's Solar Maximum Mission from the time of launch on Feb 14, 1980 to the end of the mission in Dec 1989. 12,772 events were detected, with the "vast  majority being solar flares". To assure stationarity and due to missing values during the years 1983 and 1984, we based our analysis just on the year 1982, in which 2,488 events happened. The list includes the start time, peak time, duration, and peak rate of each event. We have used "start time" as the variable for event times, and "peak rate" as the variable for event magnitudes.

@@

```{r flare-diagnostics-2, fig.height=4.5, fig.width=4.5, fig.align="center",fig.cap="\\label{fig:flare-diagnostics-2} Diagnostic plots for the solar flare data: empirical copula.",out.width="70%"}
empcopula(flares_ctre)
```

```{r flare-diagnostics-3, fig.height=4.5, fig.width=4.5, fig.align="center",fig.cap="\\label{fig:flare-diagnostics-3} Diagnostic plots for the solar flare data: QQ Plot.",out.width="70%"}
flares_ctre %>% interarrival() %>% mlqqplot(log = 'xy', tail = 0.85, main = "Mittag-Leffler QQ Plot")
```

Figure \ref{fig:flares} shows the stability plots for the solar flare data, on the left for the tail parameter and on the right for the scale parameter. Dotted lines show 95% confidence intervals, which are derived from the asymptotic normality of the log-moments estimators [@Cahoy2013] and the $\delta$-method [@MittagLeffleR], dashed lines show the deduced values of $\beta$ resp.\ $\sigma_0$. The stability plot for the tail stabilizes nicely round about 0.85 (dashed line), while it is a little bit harder to deduce an estimate for the scale parameter. Howerever, the stability plot for the scale parameter seems to stabilize round about $3 \times 10^7$ (dashed line).
